---
title: "p8131_hw7_mp3745"
author: "Matthew Parker"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(nnet)
library(MASS)
library(pscl)
library(nlme)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

#### 1.

Read in the data
```{r}
pitch_df = read_csv("./data/HW7-politeness_data.csv")
```

Provide boxplots to show the relation between gender/attitude and pitch (ignoring diô°€erent scenarios)
```{r}
pitch_df %>% 
  group_by(gender, attitude) %>% 
  ggplot(aes(x = gender, y = frequency, color = attitude)) + 
  geom_boxplot()
```

Fromt the above boxplots, it appears that females tend to have a higher frequency than males. In addition, those with informal attitude may have a slightly higher pitch than those with polite attitude.


#### 2.

Fit a mixed effects model with random intercepts for different subjects (gender and attitude being fixed effects)
```{r}
# model fit
lmm_1 = lme(frequency ~ gender + attitude, random = ~1 | subject,  data = pitch_df, method = 'REML') 

# summary
summary(lmm_1)
```

Covariance matrix for Yi
```{r}
# sigma b squared
sigma_b_sq = as.numeric(VarCorr(lmm_1)[1])

# sigma squared
sigma_sq = as.numeric(VarCorr(lmm_1)[2])

# Covariance matrix for Yi
cov_yi = matrix(
  c(sigma_b_sq + sigma_sq, sigma_b_sq, "...", sigma_b_sq, 
    sigma_b_sq, sigma_b_sq + sigma_sq, "...", sigma_b_sq, 
    "...", "...", "...", "...", 
    sigma_b_sq, sigma_b_sq, "...", sigma_b_sq + sigma_sq), 4, 4
)

cov_yi
```

Covariance matrix for REML estimates of fixed effects
```{r}
vcov(lmm_1)
```

BLUPs for subject specific intercepts
```{r}
random.effects(lmm_1)
```

Residuals
```{r}
pitch_df$frequency - fitted(lmm_1)
```


#### 3.

Fit a mixed effects model with random intercepts for different subjects (gender, attitude, and gender*attitude being fixed effects)
```{r}
# model fit
lmm_2 = lme(frequency ~ gender*attitude, random = ~1 | subject,  data = pitch_df, method = 'REML') 

# summary
summary(lmm_2)
```

Compare the two models by likelihood ratio test
```{r}
# model 1 with ML estimate
lmm_1_ml = lme(frequency ~ gender + attitude, random = ~1 | subject,  data = pitch_df, method = 'ML')

# model 2 with ML estimate
lmm_2_ml = lme(frequency ~ gender*attitude, random = ~1 | subject,  data = pitch_df, method = 'ML') 

# Compare
anova(lmm_1_ml, lmm_2_ml)
```

Since the p-value from the likelihood ratio test is > 0.05, we fail to reject the null hypothesis that the two models fit the data equally well. Because of this, it appears the interaction between gender and attitude is not significantly associated with pitch.


#### 4.


The model is handwritten.

Fit the model with random intercepts and random slopes for attitude for different subjects (gender and attitude are also fixed effects) 
```{r}
# Random intercept and slope for attitude
lmm_3 = lme(frequency ~ gender + attitude, random = ~ 1 + attitude | subject,  data = pitch_df, method = 'REML') 

# Summary
summary(lmm_3)
```

Var(Yi) is handwritten


Fixed effect, random effect, and BLUP for first female subject in scenario 1 with polite attitude
```{r}
# fixed effects
fixed.effects(lmm_3)

# random effect
ranef(lmm_3)
```

For the first female subject in scenario 1 with polite attitude, the fixed effect is 226.99, the random effect is -13.58, EBLUPs are -13.58 and -8.41e-07